Teaching Dossier

### 2020-02-24 CUT/teaching-philosophy.tex
Teaching Observations in appendices

\usepackage{apacite}
\usepackage[colorlinks = true,
linkcolor = red,
urlcolor  = blue,
citecolor = green,
filecolor = cyan,
anchorcolor = blue]{hyperref}
\usepackage{setspace}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry}
\title{Teaching Dossier}
\author{W. Justin Toth}
\begin{document}
\maketitle
\newpage
\tableofcontents
\newpage
\section{Statement of Teaching Philosophy}
\paragraph{}
Mathematics is a rich, deep tapestry of fascinating interrelated concepts about shapes and numbers. Many passionate teachers of the subject, myself included, share this sentiment. As an instructor, my goal is to transmit this perspective to my students. I want to stoke their passion, ignite their curiosity, and instill lifelong confidence in their understanding. Towards these lofty goals I have adopted methods of teaching which encourages students to take greater ownership over the material and their own learning process.

@@

\appendix
\section*{Appendices}
%Observation Reports
%Extended Abstract for Research paper
%Full evaluations
%Sample Lesson Plan from guest lecture
%Sample Assignment from CO 327
%Sample Lesson Plan from CO 327
\section{Teaching Observation Report 1}\label{sec:observation-report-1}

\section{Teaching Observation Report 2}\label{sec:observation-report-2}
\end{document}

>>>>

\usepackage{apacite}
\usepackage[colorlinks = true,
linkcolor = red,
urlcolor  = blue,
citecolor = green,
filecolor = cyan,
anchorcolor = blue]{hyperref}
\usepackage{setspace}
\usepackage{pdfpages}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry}
\title{Teaching Dossier}
\author{W. Justin Toth}
\begin{document}
\maketitle
\newpage
\tableofcontents
\newpage
\section{Statement of Teaching Philosophy}
\paragraph{}
Mathematics is a rich, deep tapestry of fascinating interrelated concepts about shapes and numbers. Many passionate teachers of the subject, myself included, share this sentiment. As an instructor, my goal is to transmit this perspective to my students. I want to stoke their passion, ignite their curiosity, and instill lifelong confidence in their understanding. Towards these lofty goals I have adopted methods of teaching which encourages students to take greater ownership over the material and their own learning process.

@@

\appendix
\section*{Appendices}
%Observation Reports
%Extended Abstract for Research paper
%Full evaluations
%Sample Lesson Plan from guest lecture
%Sample Assignment from CO 327
%Sample Lesson Plan from CO 327
\section{Teaching Observation Report 1}\label{sec:observation-report-1}
\begin{tabular}{r l}
\textbf{Event Observed:}&	CO-456: Introduction to Game Theory\\
\textbf{Date Observed:}&	November 23, 2018\\
\textbf{Location:}&	E2-1732\\
\textbf{Time:}&	11:30 am - 12:20 pm\\
\textbf{Number of Students Present:}&	39\\
\textbf{Observer:}&	Shahrukh Athar, Graduate Instructional Developer,\\
     &University of Waterloo Centre for Teaching Excellence\\
\end{tabular}

\rule{\linewidth}{0.25mm}
\textbf{Aspects to Maintain:}

\textbf{1. Use of Questions and Comprehension Checks}

Justin, it was great to see you perform comprehension checks regularly during the lesson. You carried out such checks by asking a variety of questions. You performed content-specific checks by asking open-ended questions, such as, “This polytope is always non-empty, does anyone know why?” or “Why is this condition what we want?”, while you carried out general- purpose checks by asking questions such as “At this point does anyone have any questions?”. You also asked questions that lie between the above two categories by asking questions such as “Does anyone have any questions about what correlated equilibria is?”. All of this is fantastic as it demonstrates a learner-centric approach to teaching rather than a teacher-centric one. It shows that you want to take your learners along with you as the lesson progresses and are willing to pause and clarify confusions before moving on. Since confusions at one point of the lesson may hinder learning later, it is very important to address them, especially in mathematics classes. The open-ended questions that you asked encouraged your students to think critically, which is an excellent tool for enhancing the learning experience. My only
recommendation here would be to ask you to pause a bit more after asking questions, for at least 5~10 seconds. As you pause, students would get the time to develop their answers and respond to you or think if they need to ask a question from you.

\textbf{2. Lesson Structure and Transitions}

I found your lesson’s main body to be incredibly well-structured. You had written the lesson’s outline on the board before the class started and this allowed your students to keep track of the lesson’s progress. Most courses in STEM disciplines have lessons that are composed of a number of components. Clear transitions between these components during the lesson enhances the overall structure of the lesson and brings clarity to the students in following the lesson. It was thus great to see that you clearly stated the different parts of the lesson that were being discussed at any given time. For example, after discussing correlated equilibria, you first performed a comprehension check about it and then said, “Let’s try to understand the relationship between correlated equilibria and the mixed Nash equilibria.”. Here, the comprehension check signified that one component of the lesson (discussion on correlated equilibria) was coming to an end, while the above-mentioned sentence signaled to your students that the next component of the lesson (discussion on the relationship between correlated and mixed Nash equilibria) was about to begin. I will encourage you to continue following such a structured approach in the delivery of your future lessons.

\textbf{3. Interactivity}

It is always beneficial to use a mix of instructor-learner and learner-learner interactions as we teach a lesson. This is because each of them fosters learning in a different manner. Through the use of questions, your lesson had a good amount of instructor-learner interaction throughout its length. However, what I specifically want to celebrate here is the learner-learner activity that you had towards the end of the lesson. You asked your students to solve an example (Prisoner’s dilemma) in pairs. The problem was clearly defined and was well-aligned with your learning outcomes. This activity allowed you to carry out post-assessment and provided an opportunity to your students to discuss any confusions with each other, which is great since an assignment on the topic was going to be handed out soon. I will strongly encourage you to continue using such active learning techniques in the future.\newline
\textbf{Targets for Change and Methods for Improvement:}

\textbf{1. Awareness of Learner Needs}

The lesson was conducted in a large classroom, which was deep and wide. In such classrooms, it can be difficult for students sitting further back, to see what is being written on the board or to hear the instructor. I was sitting in the last row of the classroom, and at times I had to really concentrate to see what was being written on the whiteboard as your writing size was a touch small. The doors of the classroom were left open throughout the lesson, and this was a source of distraction whenever people, who were talking to each other, passed in the adjacent corridor. In such moments, it became a bit difficult to follow what you were saying. I will encourage you to check with your students, especially those sitting at the back, as to whether your writing is clearly visible to them, and if they can hear you. Preferably such checks should be performed at the very beginning of the lesson. It may also be a good idea to use a mic in such large classrooms and to close the doors of the room at the beginning. I will also encourage you to ask your students before rubbing a section of the whiteboard. I noticed that you addressed your students as “guys” a number of times. Since there were women sitting in the classroom, they
might not have identified with the word “guys”. I will encourage you to try using terms that are more inclusive. For example, you could try “you all,” “all of you,” “you folks,” and so on.

\textbf{2. Write-Turn-Talk}

When writing on the whiteboard, it is important not to talk as your back is turned to your audience. This creates a disconnect between the teacher and the learners. For example, a student may have their hand raised to signal to the instructor that they want to ask a question, but if the instructor is not looking towards the students then they can easily miss taking that question. This also leads to the loss of eye contact, which is an important aspect of interpersonal communication as it signals interest in others and helps in classroom management (such as preventing students from getting distracted from the material). I noted that you talked while writing on the board several times during the lesson. For example, when you were discussing different cases, you were almost continuously writing on the board and talking at the same time. In your future teaching opportunities, try to incorporate the Write-Turn-Talk technique, where you should write on the board first without talking, then turn towards the learners and talk while looking at them. Doing so will enable you to better engage the students into the learning process and will help you in keeping a bird’s eye view of the class.\newline
\textbf{Additional Comments:}
Thank you for having me in this class for your lesson, Justin! It was a pleasure to observe your teaching. I wish you nothing but the best in all your future teaching endeavors.
\section{Teaching Observation Report 2}\label{sec:observation-report-2}
\begin{tabular}{r l}
    \textbf{Event Observed:}&	CO327: Deterministic OR Models\\
    \textbf{Date Observed:}&	July 16, 2019\\
    \textbf{Location:}&	RCH 206\\
    \textbf{Time:}&	10:00am – 11:00am\\
    \textbf{Number of Students Present:}&	8\\
    \textbf{Observer:}&	Caelan Wang, Graduate Educational Developer,\\
         &University of Waterloo Centre for Teaching Excellence
\end{tabular}

\rule{\linewidth}{0.25mm}
You were confident, clear, and knowledgeable in the classroom. Overall, your lesson was definitely a success.\newline
\textbf{Aspects to Maintain:}

\textbf{1. Lesson Structure and Preparation}
Justin, your lesson was certainly well-structured so that it was easy to see the main points you wanted to get to. I appreciated that you arrived a few minutes early to put the outline of the lesson on the board. It gave your learners and overview of the material covered in the lesson, which helped them to become more focused while getting prepared. Once you started teaching, you reminded your learners of the material in the previous lesson, giving them a nice warm up.
Throughout the lesson, it was evident to me that you were knowledgeable and well-prepared to teach. You always appeared to be very confident, knowing what to put on the board. If fact, you did not stumble even once in your teaching, which was impressive. Even though I didn’t get to see the end of the lesson, the structure was very clear and I had a good idea of where you wanted to get to. Well done!

\textbf{2. Handwriting and Board Use}

Your handwriting was very clear and your board work was well-organized. For example, when you were discussing the dual simplex method, I found it very helpful that you used bullet points to list out the steps of the methods. Moreover, you underlined some key words to make sure your learners know what the emphasis of the lesson were. If a learner is 10 minutes late to your lesson, they could easily get the idea of what they have missed, thanks to your great board work. With the very limited board space you had, it was particularly difficult to use the board well, but you did a great job!

The board did look a little dark from the very back of the classroom. It might be better to turn on the lights above the board and/or use coloured chalk to highlight the key words, so that they would be easier to see.

\textbf{3. Use of Examples}

I really appreciated that you used a lot of examples throughout your lesson, so that your learners could better understand the mathematical principles through practice. When you were explaining the dual simplex method, you gave an example of the linear program, and then explained the big ideas behind the dual simplex method using this example, giving your learners something concrete to work with while learning how to choose a variable to enter/leave the basis.

Throughout your teaching, you have used examples and guided questions to promote critical thinking amongst your learners. For instance, after explaining the dual simplex method, you asked how dual feasibility would be preserved. By looking at the example, your learners were able to come up with the solution to always pick the non-basic variable with the smallest index to avoid cycling. Such well-thought examples and questions really kept your learners actively engaged during your lesson. I would highly encourage you to continue these great practices in your future teaching!\\
\textbf{Targets for Change and Methods for Improvement:}

\textbf{1. Effective Questioning and Responding Techniques}

Justin, as I mentioned earlier, you have certainly asked a lot of good questions during this observation. Moreover, there was ample comprehension checks for you to make sure your learners were on track. These are all great, and I just want to provide some further tips to enhance your questioning and responding techniques.

One thing I noticed was, even though you were doing a lot of comprehension checks, your
learners weren’t always responding to them. When facing a group of relatively quiet learners, it can be hard to understand how they are doing. When you ask a question, I would recommend you to wait a little longer to give your learners enough time to form a response. I understand it can be hard to stand through the “awkward silence”, so I would recommend to
bring a bottle of water and take a sip of water after you ask a question, so that the time passes more naturally. Alternatively, if you don’t get a response after a comprehension check, sometimes I would do a poll of the class to see how they are doing.

When a learner asks a question or makes a comment, I want you to work on repeating it before further discussions. With such a small classroom, your learners could certainly hear each other. However, they would be paying attention to you instead of each other, so when a learner says something, others could easily miss it. That is why repeating what your learners said would be essential to ensure your entire classroom on the same page. I encourage you to try this out the next time you teach.

\textbf{2. Pacing, Volume, and Eye Contact}

The pace of your speech was quite fast, which might have been okay if you are teaching in a different room. However, this specific classroom did not help – it was small, but it also echoed, making it difficult to hear you clearly. I would encourage you to try to speak just a little slower so it would be easier to hear.

You asked for specific feedback on write-turn-talk, and I would say I did not see you write and talk at the same time a lot. However, when discussing the material written on the board, you had the tendency of facing the board to point at your writing. While pointing is certainly a good technique to keep your learners attention, I want you to try to be more mindful of this and try to face the learners even when you need to point at certain things on the board. When you are discussing the material, it would be particularly helpful to maintain good eye contact with your learners, so that you could make sure they are following your explanations.

\section{}
\end{document}
